---
title: "Project"
author: "Diana Rueda"
date: "3/15/2021"
output: html_document
---

```{r}
library(blblm)
library(future)
library(furrr)
library(ggplot2)
library(bench)
```



```{r}
library(tibble)
x1 = rnorm(100000, sd = .25)
x2 = rnorm(100000, sd = .25)
y = rnorm(100000, sd = .25)
X = tibble(x1,x2,y)

```

```{r}
library(RcppEigen)
```

```{r}
model_test = fastLm(y~x1*x2, data = X, weights = freqs)
```

```{r paged.print = FALSE}
result = bench::mark(
  blblm = blblm(y ~ x2 * x1, data = X, m = 5, B = 1000),
  blblm_4wkrs = blblm(y ~ x2 * x1, data = X, m = 5, B = 1000, workers = 4, fastlm = TRUE),
  relative = FALSE,
  check = FALSE,
  iterations = 10
)

```

```{r}
result
```

```{r}
test = blblm(mpg ~ wt * hp, data = mtcars, m = 3, B = 100)
```

```{r}
lm1(y, x1 , n = length(y), fastlm = FALSE )
```

```{r}
sigma(test)

```

```{r}
predict_int = predict(test, data.frame(wt = c(2.5, 3), hp = c(150, 170)), confidence = TRUE)
dim(predict_int)
```

#' @import purrr
#' @import stats
#' @importFrom magrittr %>%
#' @importFrom utils capture.output
#' @details
#' Linear Regression with Little Bag of Bootstraps
"_PACKAGE"


## quiets concerns of R CMD check re: the .'s that appear in pipelines
# from https://github.com/jennybc/googlesheets/blob/master/R/googlesheets.R
utils::globalVariables(c("."))



#' Bind two factors
#'
#' Create a new factor from two existing factors, where the new factor's levels
#' are the union of the levels of the input factors.
#'
#' @param formula formula
#' @param data dataframe
#' @param m integer
#' @param B integer
#' @param workers integer
#'
#' @return object class blblm
#' @examples
#' fit <- blblm(mpg ~ wt * hp, data = mtcars, m = 3, B = 100)
#' @export
blblm <- function(formula, data, m = 10, B = 5000, workers = 1) {
  if(workers > 1){
    suppressWarnings(plan(multiprocess, workers = 4))
    options(future.rng.onMisuse = "ignore")
    data_list <- split_data(data, m)
    estimates <- future_map(
      data_list,
      ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B))
    res <- list(estimates = estimates, formula = formula)

  }else{
    data_list <- split_data(data, m)
    estimates <- map(
      data_list,
      ~ lm_each_subsample(formula = formula, data = ., n = nrow(data), B = B))
    res <- list(estimates = estimates, formula = formula)

  }
  class(res) <- "blblm"
  invisible(res)

}

#' Bind two factors
#'
#' split data into m parts of approximated equal sizes
#'
#' @param data dataframe
#' @param m integer
#'
#' @return ataframe
#' @examples
#' split_data(data = mtcars, m = 3)
#' @export
split_data <- function(data, m) {
  set.seed(141)
  idx <- sample.int(m, nrow(data), replace = TRUE)
  data %>% split(idx)
}

#' Bind two factors
#'
#' compute the estimates
#'
#' @param formula regression formula
#' @param data dataframe
#' @param n integer
#' @param B integer
#'
#' @return list
#' @examples
#' lm_each_subsample(mpg ~ wt * hp, data = mtcars, n = 10, B = 100)
#' @export
lm_each_subsample <- function(formula, data, n, B) {
  # drop the original closure of formula,
  # otherwise the formula will pick a wrong variable from the global scope.
  environment(formula) <- environment()
  m <- model.frame(formula, data)
  X <- model.matrix(formula, m)
  y <- model.response(m)
  replicate(B, lm1(X, y, n), simplify = FALSE)
}

#' Bind two factors
#'
#' compute the regression estimates for a blb dataset
#'
#' @param X data frame
#' @param y vector
#' @param n integer
#'
#' @return vector
#' @examples
#' n <- 7 ; p <- 2
#' X <- matrix(rnorm(n * p), n, p)
#' y <- rnorm(n)
#' lm1(X, y, 10)
#' @export
lm1 <- function(X, y, n) {
  freqs <- as.vector(rmultinom(1, n, rep(1, nrow(X))))
  fit <- lm.wfit(X, y, freqs)
  list(coef = blbcoef(fit), sigma = blbsigma(fit))
}

#' Bind two factors
#'
#' compute the coefficients from fit
#' Create a new factor from two existing factors, where the new factor's levels
#' are the union of the levels of the input factors.
#'
#' @param fit fitted blblm model
#'
#' @return vector
#' @export
#' @examples
#' blbcoef(blblm(mpg ~ wt * hp, data = mtcars, m = 3, B = 100))
blbcoef <- function(fit) {
  coef(fit)
}

#' Bind two factors
#'
#' compute sigma from fit
#'
#' @param fit blblm object
#'
#' @return double
#' @examples
#' blbsigma(blblm(mpg ~ wt * hp, data = mtcars, m = 3, B = 100))
#' @export
blbsigma <- function(fit) {
  p <- fit$rank
  e <- fit$residuals
  w <- fit$weights
  sqrt(sum(w * (e^2)) / (sum(w) - p))
}


#' @export
#' @method print blblm
print.blblm <- function(x, ...) {
  cat("blblm model:", capture.output(x$formula))
  cat("\n")
}


#' @export
#' @method sigma blblm
sigma.blblm <- function(object, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  sigma <- mean(map_dbl(est, ~ mean(map_dbl(., "sigma"))))
  if (confidence) {
    alpha <- 1 - 0.95
    limits <- est %>%
      map_mean(~ quantile(map_dbl(., "sigma"), c(alpha / 2, 1 - alpha / 2))) %>%
      set_names(NULL)
    return(c(sigma = sigma, lwr = limits[1], upr = limits[2]))
  } else {
    return(sigma)
  }
}

#' @export
#' @method coef blblm
coef.blblm <- function(object, ...) {
  est <- object$estimates
  map_mean(est, ~ map_cbind(., "coef") %>% rowMeans())
}


#' @export
#' @method confint blblm
confint.blblm <- function(object, parm = NULL, level = 0.95, ...) {
  if (is.null(parm)) {
    parm <- attr(terms(object$formula), "term.labels")
  }
  alpha <- 1 - level
  est <- object$estimates
  out <- map_rbind(parm, function(p) {
    map_mean(est, ~ map_dbl(., list("coef", p)) %>% quantile(c(alpha / 2, 1 - alpha / 2)))
  })
  if (is.vector(out)) {
    out <- as.matrix(t(out))
  }
  dimnames(out)[[1]] <- parm
  out
}

#' @export
#' @method predict blblm
predict.blblm <- function(object, new_data, confidence = FALSE, level = 0.95, ...) {
  est <- object$estimates
  X <- model.matrix(reformulate(attr(terms(object$formula), "term.labels")), new_data)
  if (confidence) {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>%
      apply(1, mean_lwr_upr, level = level) %>%
      t())
  } else {
    map_mean(est, ~ map_cbind(., ~ X %*% .$coef) %>% rowMeans())
  }
}


mean_lwr_upr <- function(x, level = 0.95) {
  alpha <- 1 - level
  c(fit = mean(x), quantile(x, c(alpha / 2, 1 - alpha / 2)) %>% set_names(c("lwr", "upr")))
}

map_mean <- function(.x, .f, ...) {
  (map(.x, .f, ...) %>% reduce(`+`)) / length(.x)
}

map_cbind <- function(.x, .f, ...) {
  map(.x, .f, ...) %>% reduce(cbind)
}

map_rbind <- function(.x, .f, ...) {
  map(.x, .f, ...) %>% reduce(rbind)
}



